---
title: "Chapter 3 Exercises"
output: html_notebook
---

#### Exercise 1 (Hall of Fame Pitching Dataset)

```{r}
library(Lahman)
library(tidyverse)
library(ggplot2)
```

```{r}
hofpitching <- read_csv("../data/hofpitching.csv")
```





```{r}
hofpitching <- hofpitching %>%
  mutate(BF.group = cut(BF, 
                        c(0, 10000, 15000, 20000, 30000),
         labels = c("Less than 10000", "(10000, 15000)",
                 "(15000, 20000)", "more than 20000")))
```

```{r}
hofpitching
```

```{r}
# a)
hofpitching %>%
  group_by(BF.group) %>%
  summarize(N = n()) -> S
```

```{r}
#b)
ggplot(S, aes(BF.group, N)) + geom_col()
```

```{r}
#c)
ggplot(S, aes(BF.group, N)) +
  geom_point() + coord_flip()
```
The first graph is more effective because it is way easier to read than the one above


#### 2 (Hall of Fame Pitching Dataset (Continued))
```{r}
#a)
ggplot(hofpitching, aes(WAR)) + geom_histogram()
```

```{r}
#b)
hofpitching %>%
  # sorts by specified column
  arrange(desc(WAR)) %>%
  # handles showing the first two rows
  slice(1:2) %>%
  # selects which column to display
  select(X2, WAR)
```
The two pitchers that standout with the highest WAR are Cy Young and Walter Johnson


#### Exercise 3 (Hall of Fame Pitching Dataset (Continued))

```{r}
#a)
# mutate allows you to create a new column with a formula associated 
hofpitching <- hofpitching %>%
  mutate(WAR.Season = WAR / Yrs)

hofpitching

ggplot(hofpitching, aes(BF.group,WAR.Season)) + geom_point() + coord_flip()
```

```{r}
# b)
ggplot(hofpitching, aes(WAR.Season, BF.group)) + geom_boxplot()
```
The more batters face shows a tendency of higher WAR

#### Exercise 4 (Hall of Fame Pitching Dataset (Continued))

```{r}
hofpitching <- hofpitching %>%
  mutate(MidYear = (From + To) / 2)
hofpitching.recent <- hofpitching %>%
  filter(MidYear >= 1960)

hofpitching.recent
```

```{r}
# a)
hofpitching <- hofpitching %>%
  arrange(WAR.Season)
hofpitching
```
```{r}
# b)
ggplot(hofpitching.recent, aes(x=WAR.Season, y = WAR.Season, label=X2)) + geom_text(angel=45)
```

#### Exercise 5 (Hall of Fame Pitching Dataset (Continued))

```{r}
# a)
ggplot(hofpitching, aes(x=MidYear, y=WAR.Season)) + geom_point()
```
```{r}
# c)
ggplot(hofpitching, 
       aes(x = MidYear, y = WAR.Season, label=X2)) +
         geom_point() +
      geom_text(data = filter(hofpitching, 
              MidYear < 1900, WAR.Season < 2))
```

