---
title: "Chapter 3 Exercises"
output: html_notebook
---

#### Exercise 1 (Hall of Fame Pitching Dataset)

```{r}
library(Lahman)
library(tidyverse)
library(ggplot2)
```

```{r}
hofpitching <- read_csv("../data/hofpitching.csv")
```





```{r}
hofpitching <- hofpitching %>%
  mutate(BF.group = cut(BF, 
                        c(0, 10000, 15000, 20000, 30000),
         labels = c("Less than 10000", "(10000, 15000)",
                 "(15000, 20000)", "more than 20000")))
```

```{r}
hofpitching
```

```{r}
# a)
hofpitching %>%
  group_by(BF.group) %>%
  summarize(N = n()) -> S
```

```{r}
#b)
ggplot(S, aes(BF.group, N)) + geom_col()
```

```{r}
#c)
ggplot(S, aes(BF.group, N)) +
  geom_point() + coord_flip()
```
The first graph is more effective because it is way easier to read than the one above


#### 2 (Hall of Fame Pitching Dataset (Continued))
```{r}
#a)
ggplot(hofpitching, aes(WAR)) + geom_histogram()
```

```{r}
#b)
hofpitching %>%
  # sorts by specified column
  arrange(desc(WAR)) %>%
  # handles showing the first two rows
  slice(1:2) %>%
  # selects which column to display
  select(X2, WAR)
```
The two pitchers that standout with the highest WAR are Cy Young and Walter Johnson


#### Exercise 3 (Hall of Fame Pitching Dataset (Continued))

```{r}
#a)
# mutate allows you to create a new column with a formula associated 
hofpitching <- hofpitching %>%
  mutate(WAR.Season = WAR / Yrs)

ggplot(hofpitching, aes(BF.group,WAR.Season)) + geom_point() + coord_flip()
```

```{r}
# b)
ggplot(hofpitching, aes(WAR.Season, BF.group)) + geom_boxplot()
```
The more batters face shows a tendency of higher WAR

#### Exercise 4 (Hall of Fame Pitching Dataset (Continued))

```{r}
hofpitching <- hofpitching %>%
  mutate(MidYear = (From + To) / 2)
hofpitching.recent <- hofpitching %>%
  filter(MidYear >= 1960)

hofpitching.recent
```

```{r}
# a)
hofpitching <- hofpitching %>%
  arrange(WAR.Season)
hofpitching
```
```{r}
# b)
ggplot(hofpitching.recent, aes(x=WAR.Season, y = WAR.Season, label=X2)) + geom_text(angel=45)
```

#### Exercise 5 (Hall of Fame Pitching Dataset (Continued))

```{r}
# a)
ggplot(hofpitching, aes(x=MidYear, y=WAR.Season)) + geom_point()
```
```{r}
# c)
ggplot(hofpitching, 
       aes(x = MidYear, y = WAR.Season, label=X2)) +
         geom_point() +
      geom_text(data = filter(hofpitching, 
              MidYear < 1900, WAR.Season < 2))
```

#### 6 (Working with the Lahman Batting Dataset)
```{r}
#b)
pull_player_ids <- function(first_name, last_name){
  Master %>%
  filter(nameFirst == first_name,
         nameLast == last_name) %>%
  select(playerID) %>% pull() -> player_id
  
  return(player_id)
}

# first lets get the player ids
cobb_id <- pull_player_ids('Ty', 'Cobb')
williams_id <- pull_player_ids('Ted', 'Williams')
# there are two so only keeping first value
Master %>% filter(nameLast == "Rose", 
                  nameFirst == "Pete",
                  birthYear == 1941) %>% 
  select(playerID) %>% pull() -> rose_id

# get Batting data for these hitters
Batting %>%
  filter(playerID %in%
           c(cobb_id, williams_id, rose_id)) ->df

df
```

```{r}
# c)
get_birthyear <- function(pid){
  Master %>%
    filter(playerID == pid) %>%
    mutate(birthyear = ifelse(birthMonth >=7,
                               birthYear + 1, birthYear)) %>%
    select(playerID, birthyear)
}

bdates <- bind_rows(get_birthyear(cobb_id),
                    get_birthyear(williams_id),
                    get_birthyear(rose_id))

df %>%
  inner_join(bdates, by = "playerID") %>%
  mutate(Age = yearID - birthyear) %>%
  select(playerID, Age, H) %>%
  group_by(playerID) %>%
  mutate(CH = cumsum(H)) -> df

df
```

```{r}
# d)
ggplot(filter(df, playerID == rose_id), aes(Age, CH)) + geom_line()
```

```{r}
#e
ggplot(df, 
       aes(Age, CH, group = playerID, color = playerID)) +
  geom_line()
```

```{r}

```


#### 7 (Working with the Retrosheet Play-by-Play Dataset)

```{r}
get_player_id_retro <- function(first_name, last_name){
  Master %>%
    filter(nameLast == last_name,
           nameFirst == first_name) %>%
    select(retroID) %>% pull() -> player_id
  
  return(player_id)
}

mac_id = get_player_id_retro('Mark', 'McGwire')
sosa_id = get_player_id_retro('Sammy', 'Sosa')

fields <- read_csv('../data/fields.csv')
data1998 <- read_csv('../data/all1998.csv',
                     col_names = pull(fields, Header))

mac.data <- filter(data1998, BAT_ID==mac_id)
sosa.data <- filter(data1998, BAT_ID==sosa_id)
```


```{r}
#b) 
mac.data <-filter(mac.data, BAT_EVENT_FL == TRUE)
sosa.data <-filter(sosa.data, BAT_EVENT_FL == TRUE)
```

```{r}
#c) 
mac.data <- mutate(mac.data, PA = 1:nrow(mac.data))
sosa.data <- mutate(sosa.data, PA = 1:nrow(sosa.data))
```

```{r}
#d) 
mac.HR.PA <- mac.data %>%
  filter(EVENT_CD == 23) %>%
  pull(PA)

sosa.HR.PA <- sosa.data %>%
  filter(EVENT_CD == 23) %>%
  pull(PA)
```

```{r}
#e) 
mac.spacings <- diff(c(0, mac.HR.PA))
sosa.spacings <- diff(c(0, sosa.HR.PA))
```


```{r}
HR_Spacing <- rbind(data.frame(Player="McGwire",
                               Spacing=mac.spacings),
                    data.frame(Player="Sosa",
                               Spacing=sosa.spacings))
```


```{r}
ggplot(HR_Spacing, aes(Spacing)) +
  geom_histogram() + facet_wrap(~ Player, ncol = 1)
```

```{r}
HR_Spacing %>% group_by(Player) %>% 
  summarize(M = median(Spacing))
```


